Homicide Increases in the COVID-19 Era

### 2025-03-01 Draft2.qmd
Update Draft2.qmd

>>>>

## Time series analyzing homicide rate trends over time by socioeconomic quartiles

```{r, echo=FALSE}
# Filter only homicide cases
homicides <- crime_la %>%
  inner_join(crimes_with_zips, by = "crime_id") %>%  
  filter(offense_type == "HOMICIDE") %>%
  mutate(date = as.Date(date))

# Join homicide data with income data
homicides <- homicides %>%
  left_join(cali_median_income, by = c("zip_code" = "ZCTA"))

# Convert median income to numeric if needed
homicides$median_income <- as.numeric(homicides$median_income)

# Create income quartiles
homicides <- homicides %>%
  mutate(income_quartile = ntile(median_income, 4))  # Divides into 4 quartiles

# Aggregate homicide counts by year and income quartile
homicide_trends <- homicides %>%
  mutate(year = lubridate::year(date)) %>%
  group_by(year, income_quartile) %>%
  summarize(homicide_count = n(), .groups = "drop")

# Plot homicide trends over time by income quartile
ggplot(homicide_trends, aes(x = year, y = homicide_count, color = factor(income_quartile))) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Homicide Rate Trends Over Time by Socioeconomic Quartiles",
    x = "Year",
    y = "Number of Homicides",
    color = "Income Quartile"
  ) +
  theme_minimal() +
  theme(
    legend.position = "top",
    text = element_text(size = 14)
  )
```
